---
title: "STAT 3340 - Final Project"
author: "Yurunyun Wang & Yuqiao Du & Rao Huang & Ruoxin Xu"
date: "12/1/2020"
output: pdf_document
---

```{r}
#install packages
install.packages("ggplot2")
install.packages("olsrr")
install.packages("dplyr")
```

```{r}
library(rlang)
library(readr)
library(ggplot2)
library(dplyr)
```
The meaning of each column is listed below:
age: age of primary beneficiary

sex: insurance contractor gender, female, male

bmi: Body mass index, providing an understanding of body, weights that are relatively high or low relative to height,
objective index of body weight (kg / m ^ 2) using the ratio of height to weight, ideally 18.5 to 24.9

children: Number of children covered by health insurance / Number of dependents

smoker: Smoking

region: the beneficiary's residential area in the US, northeast, southeast, southwest, northwest.

charges: Individual medical costs billed by health insurance
```{r}
#import the data from a csv file and add new data point
datasetfull <- read_csv("insurance.csv")
# one new additional data point
datasetfull[nrow(datasetfull) + 1,] <- list(22,"female",23.71,0,"no","northeast",5503.7768)
```

```{r}
#count the categorical data
count(datasetfull, sex)
count(datasetfull, region)
count(datasetfull, smoker)
```

```{r}
#calculate the correlation matrix
keep<-c('age','bmi','children','charges')
num_data<-subset(datasetfull, select=keep)
cor(num_data)
```

```{r}
#draw the scatterplot
pairs(~age+bmi+children+charges,data=datasetfull)
```

```{r}
#see the summary imformation of data
summary(datasetfull)
```


```{r}
#draw ggplot for categorical data
ggplot(data = datasetfull, aes(x = sex, y = charges)) + geom_boxplot() + labs(title = "Charges distribution based on sex")
```

```{r}
ggplot(data = datasetfull, aes(x = as.factor(children), y = charges)) + geom_boxplot() + labs(x = "Number of Children") + labs(title = "Charges distribution based on Number of Children")
```

```{r}
ggplot(data = datasetfull, aes(x = smoker, y = charges)) + geom_boxplot() + labs(title = "Charges distribution based on smoker")
```

```{r}
ggplot(data = datasetfull, aes(x = region, y = charges)) + geom_boxplot() + labs(title = "Charges distribution based on region")
```
```{r}
bmi_category <- cut(datasetfull$bmi, c(0, 18.5, 24.9, 29.9, 10000), labels = c("Underweight","Healthy weight","Overweight","Obese"), right = FALSE)
bmi_category <- data.frame((bmi_category))
bmidata <- cbind(charge = datasetfull$charges, bmi = bmi_category)
colnames(bmidata) <- c("charge", "bmi")
ggplot(data = bmidata, aes(x = bmi, y = charge)) + geom_boxplot() + labs(title = "Charges distribution based on bmi class")
```

```{r}
age_range <- data.frame(ifelse(datasetfull$age < 20, "Minor", "Adult"))
agedata <- cbind(datasetfull$charges, age_range)
colnames(agedata) <- c("charge", "age")
ggplot(data = agedata, aes(x = age, y = charge)) + geom_boxplot() + labs(title = "Charges distribution based on age")
```

```{r}
  ggplot(agedata, aes(x = age))+geom_histogram(stat="count", fill = "white", color = "grey") + labs(title = "Age distribution")
```

```{r}
#set indicator variable
datasetfull$sex<-as.factor(datasetfull$sex)
datasetfull$region<-as.factor(datasetfull$region)
datasetfull$smoker<-as.factor(datasetfull$smoker)
datasetfull$sex_male <- ifelse(datasetfull$sex == "male", 1, 0)
datasetfull$sex_female <- ifelse(datasetfull$sex == "female", 1, 0)
datasetfull$region_southwest <- ifelse(datasetfull$region == "southwest", 1, 0)
datasetfull$region_southeast <- ifelse(datasetfull$region == "southeast", 1, 0)
datasetfull$region_northwest <- ifelse(datasetfull$region == "northwest", 1, 0)
datasetfull$region_northeast <- ifelse(datasetfull$region == "northeast", 1, 0)
datasetfull$smoker_yes <- ifelse(datasetfull$smoker == "yes", 1, 0)
datasetfull$smoker_no <- ifelse(datasetfull$smoker == "no", 1, 0)
datasetfull <- subset(datasetfull, select = -c(sex, region, smoker))
```

```{r}
#the analysis of model without interactions and use backward to obtain the 'best model'
fullmodel1=lm(charges~region_southwest+region_southeast+region_northwest+region_northeast+age+bmi+children+sex_male+smoker_yes+sex_female+smoker_no,data=datasetfull)
step1=step(fullmodel1,direction="backward")
summary(step1)
```

#the analysis of model with interactions and use backward to obtain the 'best model'
```{r}
#the analysis of model with interactions and use backward to obtain the 'best model'
fullmodel2=lm(charges~region_southwest+region_southeast+region_northwest+region_northeast+age+bmi+children+sex_male*smoker_yes+sex_female*smoker_no+sex_female*smoker_no+sex_male*smoker_yes,data=datasetfull)
step2=step(fullmodel2,direction="backward")
summary(step2)
```

#estimate the chosen model
```{r}
#estimate the chosen model
anova(step1,step2)
```
```{r}
anova(step2)
```
```{r}
plot(step2)
# Residuals vs Fitted:
# Looking at the Residuals vs Fitted plot, we can see that the red line is approximately flat, which means that the mean of error is approximately equal to 0.
# Normal Q-Q:
# The residuals are approximately matched to the diagonal line, which means that the residuals are roughly normally distributed. And we can see that both the upper and lower tail are ‘heavier’ (have larger values) than what we would expect under the gauss-markov assumptions. 
# Scale-Location:
# Looking at the Scale-Location plot, we can see that the red line is approximately flat, which means that there is no indication of having non-constant variance.
# Residuals vs Leverage:
# The Residuals vs Leverage plot indicate that there is no evidence of outliers because the “Cook’s distance” dashed curves don’t even appear on the plot. And there exists some leverage points: 1301, 578, 544.
```

#check whether the variables are independent
```{r}
install.packages("MPV")
install.packages("faraway")
library(MPV)
library(faraway)
vif(step2)
# As for the multicolinearity, since all the attributes' vif are less than 10, we can conclude that these variables are independent.
```